\Title

### 2023-02-09 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.
method first draft

\subsection{Task definition: fast motion detection}

@@

From the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation. Indeed, by discretization of time, we may transform any event-based input as the boolean matrix $A \in \{0, 1 \}^{N\times T}$, where $N$ is the number of neurons in $\presynaddrspace$ and $T$ is the number of time bins. In this simplified model, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ giving for each neuron $\postsynaddr$ the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in \synapse^\postsynaddr}, \kernel^\postsynaddr(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) = \synapticweight_\ranksyn$. Therefore, if we make the approximation that one neuron integrates only on the temporal window given by its variety of synaptic delays, the integration of the spike train as input can be formalized by a 3D convolution operation. The longest synaptic delay defines the depth of the kernel $\kernel^\postsynaddr$ and all possible delays associated to the different presynaptic addresses are represented. Using this dense representation, the integration of the spike train as input by the synapses of neuron $\postsynaddr$ is:

@@

@@

\note{describe the learning -> analogy with hebbian learning from TPAMI}

\begin{figure}%[h!]%[ht!]
    \centering
    \includegraphics[width=\linewidth]{figures/HD-SNN.png}
    \caption{
    \textbf{- HD-SNN illustration. From left to right:} A 3D representation of the event stream as output of the DVS. A 3D convolution is applied, on the dense representation of the input, along the $X$ and $Y$ spatial axes (defining the DVS pixel grid) and the temporal axis. There are as many kernels as the possible number of motion classes in the recording. For clarity, only $4$ kernels are illustrated and some weights are not represented in the voxel grid. For each voxel on the dense representation of the event stream as input, the output of the 3D convolution is processed by the non-linearity of the MLR model (i.e. softmax function). The output of the MLR gives a probability for each class, associated to a specific kernel, to emit an output event, or spike. With an appropriate spiking mechanism, here a winner-takes-all, we obtain, as output of the HD-SNN model, a new spike train with the different spikes associated to a specific motion class. }
    \label{fig:model}
\end{figure}
%
The output of the MLR model corresponds to an event with the highest probability class, keeping the same timing as the input event. It is important to note that we include the prior knowledge that the event-based input is spatiotemporally clustered and that these clusters are sparse. Therefore, we will only select a given proportion of the most active cells, which we set here to $.01\%$ of all output voxels. This threshold was found using cross-validation and future work should automate how this may be determined for various datasets. Given a stimulus input, the model thus yields a spike output in the postsynaptic address space.

@@

Since the model is fully differentiable, we can now implement a semi-supervised learning rule. The loss function of the MLR model is the binary cross-entropy on the output of the classification layer that can be extended to space-time convolutions. Supervision was implemented using the input binary events as defined above and the labels as the desired output. The labels were defined at each time point as a one-shot encoding of the current motion in the channel corresponding to the current motion for all positions. As in this semi-supervised context, the label is known, but the timing is not, we used a selection of the temporal support in an unsupervised manner. It is important to note that we weighted the cost function by considering only active cells, such that the error is only back propagated to the spatial locations of these most active cells. This is reminiscent of previous methods solving this problem using a Winner-Takes-All mechanism~\citep{masquelier_unsupervised_2007}. Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$). We tested the effect of different parameters on a validation set to quantify the role of each parameter.

@@

Finally, the output of the MLR model is an event-based representation predicting at each position and at each time the probability of each motion. Such an output gives a form of optical flow that can be exploited for non-rigid motions, but we have defined here, for simplicity, an evaluation method that applies to our task with a full field motion. We have shown above that when different independent observations (here, the estimated motion at different spatial locations) are recognized as having a common cause (here, the rigid motion of the image), then an optimal estimate of the logit of this probability is the sum of the logits of the independent probabilities. By taking the logit of the probability of the output given by the model, we can therefore calculate the probability of the output. This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted) or the squared error % $|| \speed - ||^2$ 
of the velocity estimate given by the average estimate. 
%$\hat{\speed} = \sum_{\class} \speed_\class \cdot Pr(\class \vert image)$ 
These calculations are performed on a different input dataset than the one used in the training or validation steps. These metrics will be useful to compare our methods with other methods on our defined task.  The complete code to reproduce the results of this paper is available at \url{https://github.com/SpikeAI/pyTERtorch}.  

>>>>

\subsection{Task definition: fast motion detection}\label{sec:task}

@@

From the perspective of simulating such event-based computations on standard CPU- or GPU-based computers and then using parallel computing, it is useful to transform this event-based representation into a dense representation. Indeed, by discretization of time, we may transform any event-based input as the boolean matrix $A \in \{0, 1 \}^{N\times T}$, where $N$ is the number of neurons in $\presynaddrspace$ and $T$ is the number of time bins. In this simplified model, we will consider that heterogeneous delays are limited in range such that the synaptic set can be represented by the dense matrix $\kernel^\postsynaddr$ of size $(N_{tb}, N_x, N_y)$, where $N_{tb}$ is the number of time bins, $N_x$ and $N_y$ are the number of pixels in both spatial dimensions. For each neuron $\postsynaddr$, $\kernel^\postsynaddr$ gives the weights as a function of presynaptic address and delay: $\forall {\ranksyn \in \synapse^\postsynaddr}, \kernel^\postsynaddr(\presynaddr_\ranksyn, \synapticdelay_\ranksyn) = \synapticweight_\ranksyn$. Therefore, if we make the approximation that one neuron integrates only on the temporal window given by its variety of synaptic delays, the integration of the spike train as input can be formalized by a 3D convolution operation. The longest synaptic delay defines the depth of the kernel $\kernel^\postsynaddr$ and all possible delays associated to the different presynaptic addresses are represented. Using this dense representation, the integration of the spike train as input by the synapses of neuron $\postsynaddr$ is:

@@

Once explained the general framework used to solve the ecological task described in section~\ref{sec:task}, we can add some heuristics based on neuroscientific observations to constrain our model and its strategies to solve the task. We set \note{arbitrarily} the size of the kernels to $(9, 17, 17)$ and define as many classes as the number of motion direction: $N_\class = 8$. Because receptive fields of biological neurons are mainly circular and can be more or less elongated, we apply a circular mask on the spatial dimensions of the kernels. And because the synaptic delay is linked, by physical principles, to the distance between the presynaptic neuron and the postsynaptic one, we apply a linear mask on the temporal axis such that the connectivity depends on the delay: the lower the delay the lower the circular mask radius (see Figure~\ref{fig:kernels} for an illustration). The second heuristics is based on the observation that biological neurons have a sparse spiking activity. From this observation, we set an output firing rate for the layer of spiking neurons in order to make a sparse selection of the events at which the model makes a decision. It results in the selection of the coefficients with the maximum likelihood value. Therefore, we will only select a given proportion of the most active cells, which we set here to $.01\%$ of all output voxels. This threshold was found using cross-validation and future work should automate how this may be determined for various datasets. Given a stimulus input, the model thus yields a spike output in the postsynaptic address space.

@@

    \includegraphics[width=\linewidth]{figures/HD-SNN.png}
    \textbf{- HD-SNN illustration. From left to right:} A 3D representation of the event stream as output of the DVS. A 3D convolution is applied, on the dense representation of the input, along the $X$ and $Y$ spatial axes (defining the DVS pixel grid) and the temporal axis. There are as many kernels as the possible number of motion classes in the recording. For clarity, only $4$ kernels are illustrated and some weights are not represented in the voxel grid. For each voxel on the dense representation of the event stream as input, the output of the 3D convolution is processed by the non-linearity of the MLR model (i.e. softmax function). The output of the MLR gives a probability for each class, associated to a specific kernel, to emit an output event, or spike. With an appropriate spiking mechanism, here a winner-takes-all, we obtain, as output of the HD-SNN model, a new spike train with the different spikes associated to a specific motion class. }
    \label{fig:model}

@@

Since the model is fully differentiable, we can now implement a semi-supervised learning rule. The loss function of the MLR model is the binary cross-entropy on the output of the classification layer. Supervision was implemented using the input binary events as defined above and the motion direction labels as the desired output. The labels were defined at each time point as a one-shot encoding of the current motion in the channel corresponding to the current motion for all positions. As in this semi-supervised context, the label is known, but the timing is not, we used a selection of the temporal support in an unsupervised manner. It is important to note that we weighted the cost function by considering only active cells, such that the error is only back propagated to the spatial locations of these most active cells. This is reminiscent of previous methods solving this problem using a Winner-Takes-All mechanism~\citep{masquelier_unsupervised_2007}. Simulations are performed with the PyTorch library using gradient descent with Adam (for $2^{12}$ movies and a learning rate of $10^{-5}$). We demonstrated in a previous work that this type of learning algorithm can be assimilated to a Hebbian learning mechanism~\cite{grimaldi_robust_2022}.  

@@

Finally, the output of the MLR model is an event-based representation predicting at each position and at each time the probability of each motion. Such an output gives a form of optical flow that can be exploited for non-rigid motions, but we have defined here, for simplicity, an evaluation method that applies to our task with a full field motion. We have shown above that when different independent observations (here, the estimated motion at different spatial locations) are recognized as having a common cause (here, the rigid motion of the image), then an optimal estimate of the logit of this probability is the sum of the logits of the independent probabilities. By taking the logit of the probability of the output given by the model, we can therefore calculate the probability of the output. This allows one to calculate the accuracy (as the percentage of times the motion is accurately predicted) or the squared error of the velocity estimate given by the average estimate. 
These calculations are performed on a different input dataset than the one used in the training or validation steps. The complete code to reproduce the results of this paper is available at \url{https://github.com/SpikeAI/pyTERtorch}.  \note{à rendre publique}
